---
title: "gt"
subtitle: "a grammar of tables"
author: "Gina Reynolds, April 2020"
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
---




class: left, top, inverse
background-image: url(https://images.unsplash.com/photo-1553321846-ad6616f5d1db?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1052&q=80)
background-size: cover

# .large[{gt} grammar of tables]
### An exploration <br>with {flipbookr}<br>and {xaringan}

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

#### .right[[gt.rstudio.com](gt.rstudio.com)<br>Edited: Gina Reynolds<br>Photo Credit: Regina Victorica in Mexico City, Mexico]


```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = T, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```

---


`r chunk_reveal("gt")`


```{r gt, include = F}
library(gt)
exibble %>% 
  rename(fruit = char) %>%
  rename(team = fctr) %>% 
  rename(weight = num) %>% 
  rename(holiday = date) %>% 
  rename(schedule = time) %>% 
  rename(appointment = datetime) %>% 
  rename(price = currency) %>% 
  rename(id = row) %>% 
  rename(club = group) ->
example
example %>% 
  .[1:5, 1:4] %>% 
  gt() %>%
  fmt_number(columns = vars(weight), decimals = 2) %>%
  fmt_date(columns = vars(holiday), date_style = 1) 
```


---

`r chunk_reveal("gt2")`


```{r gt2, include = F}
library(gt)
example %>%
  .[1:5, 5:8] %>% 
  gt() %>%
  fmt_time(columns = vars(schedule), time_style = 1) %>%
  fmt_datetime(columns = vars(appointment), 
               date_style = 6, 
               time_style = 4) %>%
  fmt_currency(columns = vars(price), 
               currency = "EUR") %>%
  tab_header(
    title = md("This is the `exibble` dataset in **gt**"), 
    subtitle = "It is one of six datasets in the package"
  ) %>%
  tab_source_note(source_note = 
                    md("More information is available at `?exibble`.")) %>% 
  tab_spanner(label = "Dates and Times", 
              columns = matches("schedule|appointment"))
```

---
class: middle, inverse, center

# Five time formats



---

`r chunk_reveal("time", break_type = "rotate")`

```{r time, include = F}
library(gt)  
example %>%  
  .[1:5, 5] %>%  
  gt() %>%  
  fmt_time(columns = vars(schedule), time_style = "hms") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "hm") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "hms_p") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "hm_p") %>% #ROTATE
  fmt_time(columns = vars(schedule), time_style = "h_p") #ROTATE
```


---
class: middle, inverse, center

# Fourteen date formats


---

`r chunk_reveal("date", break_type = "rotate")`

```{r date, include = F}
library(gt)  
example %>%  
  .[1:5, 4] %>%  
  gt() %>%  
  fmt_date(columns = vars(holiday), date_style = "iso") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "wday_month_day_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "wd_m_day_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "wday_day_month_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "month_day_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "m_day_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "day_m_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "day_month_year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "day_month") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "year") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "month") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "day") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "year.mn.day") %>% #ROTATE
  fmt_date(columns = vars(holiday), date_style = "y.mn.day") #ROTATE
```




---
class: middle, inverse, center

# *Many* currencies


---

`r chunk_reveal("currency", break_type = "rotate")`

```{r currency, include = F}
library(gt)  
example %>%  
  .[1:8, 7] %>%  
  gt() %>%  
  fmt_currency(columns = vars(price), currency = "AED") %>% # United Arab Emirates dirham  #ROTATE
  fmt_currency(columns = vars(price), currency = "AFN") %>% # Afghan afghani  #ROTATE
  fmt_currency(columns = vars(price), currency = "BAM") %>% # Bosnia and Herzegovina convertible mark  #ROTATE
  fmt_currency(columns = vars(price), currency = "BBD") %>% # Barbados dollar  #ROTATE
  fmt_currency(columns = vars(price), currency = "BDT") %>% # Bangladeshi taka  #ROTATE
  fmt_currency(columns = vars(price), currency = "BGN") %>% # Bulgarian lev  #ROTATE
  fmt_currency(columns = vars(price), currency = "BHD") %>% # Bahraini dinar  #ROTATE
  fmt_currency(columns = vars(price), currency = "TMT") %>% # Turkmenistani manat  #ROTATE
  fmt_currency(columns = vars(price), currency = "TND") %>% # Tunisian dinar  #ROTATE
  fmt_currency(columns = vars(price), currency = "TOP") %>% # Tongan pa'anga  #ROTATE
  fmt_currency(columns = vars(price), currency = "USD") %>% # United States dollar  #ROTATE
  fmt_currency(columns = vars(price), currency = "USN") %>% # United States dollar  #ROTATE
  fmt_currency(columns = vars(price), currency = "USS") %>% # United States dollar  #ROTATE
  fmt_currency(columns = vars(price), currency = "UYI") %>% # Uruguay Peso en Unidades Indexadas  #ROTATE
  fmt_currency(columns = vars(price), currency = "UYU") %>% # Uruguayan peso  #ROTATE
  fmt_currency(columns = vars(price), currency = "UZS") %>% # Uzbekistan som  #ROTATE
  fmt_currency(columns = vars(price), currency = "VEF") %>% # Venezuelan bolivar  #ROTATE
  fmt_currency(columns = vars(price), currency = "VND") %>% # Vietnamese dong  #ROTATE
  fmt_currency(columns = vars(price), currency = "VUV") %>% # Vanuatu vatu  #ROTATE
  fmt_currency(columns = vars(price), currency = "XPD") %>% # Palladium  #ROTATE
  fmt_currency(columns = vars(price), currency = "XPF") %>% # CFP franc  #ROTATE
  fmt_currency(columns = vars(price), currency = "XPT") %>% # Platinum  #ROTATE
  fmt_currency(columns = vars(price), currency = "XSU") %>% # XSU  #ROTATE
  fmt_currency(columns = vars(price), currency = "XUA") %>% # XUA  #ROTATE
  fmt_currency(columns = vars(price), currency = "YER") %>% # Yemeni rial  #ROTATE
  fmt_currency(columns = vars(price), currency = "ZAR") %>% # South African rand  #ROTATE
  fmt_currency(columns = vars(price), currency = "ZMW") # Zambian kwacha  #ROTATE
```



---

`r chunk_reveal("pizza")`

```{r pizza, include = F}
library(tidyverse)
library(paletteer)
library(gt)
pizzaplace %>%
  mutate(type = case_when(
    type == "chicken" ~ "chicken (pizzas with chicken as a major ingredient)",
    type == "classic" ~ "classic (classical pizzas)",
    type == "supreme" ~ "supreme (pizzas that try a little harder)",
    type == "veggie" ~ "chicken (pizzas without any meats whatsoever)",
  )) %>%
  mutate(size = factor(size, levels = c("S", "M", "L", "XL", "XXL"))) %>%
  dplyr::group_by(type, size) %>%
  dplyr::summarize(
    sold = n(),
    income = sum(price)
  ) %>% 
  # table is kind of large, hard to make sense of
  dplyr::filter(size != "S" & size != "M" & size != "XL") ->
pizza_prep
```


---

`r chunk_reveal("pizza_prep")`

```{r pizza_prep, include = F}
pizza_prep %>% 
  gt(rowname_col = "size") %>%
  tab_header(title = md("&#127829; Pizzas Sold in 2015 &#127829;")) %>%
  fmt_number(
    columns = vars(sold),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_currency(
    columns = vars(income),
    currency = "USD"
  ) %>%
  cols_align(align = "right", columns = TRUE) ->
pizza2
```

---

`r chunk_reveal("pizza1")`

```{r pizza1, include = F}
pizza2 %>% 
  data_color(
    columns = vars(sold, income),
    colors = scales::col_numeric(
      palette = paletteer::paletteer_d(
        palette = "ggsci::red_material"
      ) %>% as.character(),
      domain = NULL
    ),
    alpha = 0.8
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(sold),
    fns = list(TOTAL = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(income),
    fns = list(TOTAL = "sum"),
    formatter = fmt_currency,
    currency = "USD"
  ) %>%
  grand_summary_rows(
    columns = vars(sold),
    fns = list(`GRAND TOTAL` = "sum"),
    formatter = fmt_number,
    decimals = 0,
    use_seps = TRUE
  )  ->
pizza3
```

---

`r chunk_reveal("pizza2")`

```{r pizza2, include = F}
pizza3 %>%
  grand_summary_rows(
    columns = vars(income),
    fns = list(`GRAND TOTAL` = "sum"),
    formatter = fmt_currency,
    currency = "USD"
  ) %>%
  tab_options(table.font.size = "tiny") %>% 
  tab_footnote(
    footnote = "The pizza category with the highest total sales.",
    locations = cells_row_groups("classic (classical pizzas)")
  ) %>%
  tab_footnote(
    footnote = md("Custom sizes for **The Greek** pizza."),
    locations = cells_stub(c("XL", "XXL"))
  ) %>%
  tab_footnote(
    footnote = md("This is a new record. Truly, 2015 was a **great** year for the `pizzaplace`."),
    locations = cells_grand_summary(columns = vars(sold))
  ) ->
pizza_tab  
```  
  
  




```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 60%}
.left-code {
  color: #777;
  width: 38%;
  height: 92%;
  float: left;
}
.right-output {
  width: 60%;
  float: right;
  padding-left: 1%;
}


@media print {
  .has-continuation {
    display: block;
  }
}
```

```
