---
title: "long to wide or wide to long"
author: "Evangeline Reynolds"
date: "2/16/2019"
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


https://images.unsplash.com/photo-1590605103416-230704277b05?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1500&q=80


class: left, top, inverse
background-image: url(https://images.unsplash.com/photo-1590605103416-230704277b05?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1500&q=80)
background-size: cover

# .large[Reshaping (Long <--> Wide)]
### Walk-throughs <br>with {flipbookr}<br>and {xaringan}

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

#### .right[Gina Reynolds<br>Photo Credit: Courtney Cook]


---

```{r}
library(flipbookr)
knitr::opts_chunk$set(cache = F, comment = "")
```

---
name: longtowideandback
## Exercise: *spread data, gather data*

Sometimes, we need to transform the shape of our data.  The wide form of data is an efficient way to store data - especially time-series data.  However, we often need data in a long for for data analysis, so shape transformations are important.  



---
## Exercise: *long to wide*

---

```{r long_to_wide, include = F}
library(gapminder)
library(tidyverse)
gapminder %>% 
  select(country, continent, 
         lifeExp, year) %>% 
  pivot_wider(names_from = year, 
              values_from = lifeExp) ->
gapminder_life_exp_wide 
```


`r chunk_reveal("long_to_wide",widths = c(49, 50))`

---
## Exercise: *wide to long*

---


```{r wide_to_long, include = F}
gapminder_life_exp_wide %>% 
  pivot_longer(`1952`:`2007`, # backticks needed because numeric
               # column names look like number values without them
               names_to = "year", 
               values_to = "lifeExp") %>% 
  mutate(year = as.numeric(year)) ->
  gapminder_life_exp_long 
```


`r chunk_reveal("wide_to_long",widths = c(49, 50))`

```{css, include = F}
@media print {
  .has-continuation {
    display: block;
  }
}
```







