---
title: "Introduction to Data Wrangling"
author: "Evangeline Reynolds"
date: "4/19/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Answer the following questions

### Q1

Execute the code and answer the questions.  Try to make a guess about the answer before you excute the code


```{r}
1:10 == 4
```

How many TRUEs are returned above?

> your answer


### Q2

```{r}
1:10 != 4
```

How many TRUEs are returned above?

> your answer




### Q3

```{r}
cars %>% 
  slice(1:5)
```

How many rows of data are returned above?

> your answer




### Q4

```{r}
cars %>% 
  slice(2:10)
```

*How many rows* of data are returned above?

> your answer



### Q5

```{r}
cars %>% 
  filter(dist < 5)
```


*List the values* in the column distance returned above?

> your answer




### Q6

```{r}
library(gapminder)
gapminder %>% 
  filter(country == "Peru")
```


*How many rows* are returned upon executing the above code?

> your answer




### Q7

```{r}
gapminder %>% 
  filter(country == "Peru") %>% 
  filter(year == 1992)
```


*How many rows* are returned upon executing the above code?

> your answer





### Q8

```{r}
gapminder %>% 
  select(country, year, lifeExp) 
```


*How many columns* are returned upon executing the above code?

> your answer





### Q9

```{r}
gapminder %>% 
  select(1:4) 
```


*How many columns* are returned upon executing the above code?

> your answer





### Q10

```{r}
gapminder %>% 
  select(country, year, lifeExp) %>% 
  rename(life_exp = lifeExp)
```


*What are the names of the columns* returned upon executing the above code?

> your answer





### Q11


```{r}
gapminder %>% 
  select(country, year, gdpPercap, pop) %>% 
  mutate(my_new_var = gdpPercap * pop)
```


*What are the names of the columns* returned upon executing the above code?

> your answer

*What might be a better name for my_new_var?*

> your answer






### Q12

```{r}
gapminder %>% 
  arrange(lifeExp)
```


*What is the first value in the column lifeExp* in the returned data frame upon executing the above code? *Which country is this associated with? In what year?*

> your answer






### Q13

```{r}
gapminder %>% 
  arrange(-lifeExp)
```


*What is the first value in the column lifeExp* in the returned data frame upon executing the above code? *Which country is this associated with? In what year?*

> your answer





### Q14

```{r}
gapminder %>% 
  distinct(continent)
```

*How many rows are there* in the returned data frame upon executing the above code?

> your answer






### Q15


```{r}
gapminder %>% 
  summarize(mean_life_exp = mean(lifeExp))
```

*How many means are reported* in the returned data frame upon executing the above code?

> your answer






### Q16

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarize(mean_life_exp = mean(lifeExp))
```

*How many means are reported* in the returned data frame upon executing the above code?

> your answer






### Q17

```{r}
gapminder %>% 
  filter(year == 2007) %>% 
  select(-year) %>% 
  mutate(europe = continent == "Europe") %>% 
  mutate(europe_category =
           case_when(europe == T ~ "Europe",
                     europe == F ~ "Not Europe")) 
```

What are the values you observe in the columns named europe and europe_cat?

> your answer

