---
title: "Data manipulation practice"
author: "Evangeline Reynolds"
date: ""
output:
  xaringan::moon_reader:
    seal: false
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



class: left, bottom, inverse
background-image: url(https://images.unsplash.com/photo-1533179856364-b2a4a66fa83e?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1267&q=80)
background-size: 87%
background-position: 40% 40%

# .column.Large[Wrangle]
## April 23, 2020
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>



#### .column[Gina Reynolds<br>  Photo Credit: Joshua Coleman]


```{r include = F}
knitr::opts_chunk$set(echo = F, cache = T, message = F, warning = F, comment = "")
i = 1
library(flipbookr)
```

---
class: inverse, middle center

# Set up

---

The following exercises are designed to help you practice your data wrangling "vocabulary". In the exercises, you'll always start from the full gapminder data and wrangle to a target using functions fromm {dplyr} (one of the packages made available in the tidyverse).

--

Load the following packages:

```{r, echo = T}
library(gapminder)
library(tidyverse)
```

--


Make sure to that the number of observations and columns in your result match the target outcome.  Let's recall the full dimensions of the gapminder data: 

```{r, echo = T}
# dim gives number of rows, then nubmer of columns
dim(gapminder)
```

--

Some other features to have in mind:

- {gapminder} only reports data every 5 years, from 1952 to 2007
- only {142} countries are included in the data


---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("selection")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("selection", display_type = "output")`

```{r selection, include = F}
gapminder %>% 
  select(country, year, lifeExp)
```

---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`


```{r, echo = F, eval = T, code = knitr::knit_code$get("filter")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("filter", display_type = "output")`

```{r filter, include = F}
gapminder %>% 
  filter(year == 2002)
```

---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`


```{r, echo = F, eval = T, code = knitr::knit_code$get("filterselect")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("filterselect", display_type = "output")`

```{r filterselect, include = F}
gapminder  %>% 
  filter(country == "Belgium")
```


---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`


```{r, echo = F, eval = T, code = knitr::knit_code$get("filterarrselect")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("filterarrselect", display_type = "output")`

```{r filterarrselect, include = F}
gapminder %>% 
  select(country, year, pop) %>% 
  filter(pop > 1000000000) %>% 
  arrange(-pop) 
```


---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`


```{r, echo = F, eval = T, code = knitr::knit_code$get("filterage")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("filterage", display_type = "output")`

```{r filterage, include = F}
gapminder %>% 
  select(country, lifeExp, year) %>% 
  filter(lifeExp > 81) %>% 
  arrange(-lifeExp)
```




---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`


```{r, echo = F, eval = T, code = knitr::knit_code$get("foursteps")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("foursteps", display_type = "output")`

```{r foursteps, include = F}
gapminder %>% 
  mutate(gdp = pop * gdpPercap) %>% 
  filter(year == 1952) %>% 
  filter(continent == "Oceania") %>% 
  select(country, gdp, year)
```


---
class: center, inverse, middle

# Wrangle to it. #`r i`


---

## Target #`r i` `r i <- i+1`


```{r, echo = F, eval = T, code = knitr::knit_code$get("indicator")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("indicator", display_type = "output")`

```{r indicator, include = F}
gapminder %>% 
  mutate(asia = continent == "Asia")
```




---
class: center, inverse, middle

# Summarizing (collapsing across rows)




---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("collapse")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("collapse", display_type = "output")`

```{r collapse, include = F}
gapminder %>% 
  distinct(continent)
```





---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("count")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("count", display_type = "output")`

```{r count, include = F}
gapminder %>% 
  filter(year == 2002) %>% 
  count(continent)
```





---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("distinctyear")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("distinctyear", display_type = "output")`

```{r distinctyear, include = F}
gapminder %>% 
  distinct(year)
```





---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("mean")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("mean", display_type = "output")`

```{r mean, include = F}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(mean_life_exp = mean(lifeExp))
```




---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("countries")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("countries", display_type = "output")`

```{r countries, include = F}
gapminder %>% 
  distinct(country)
```




---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("median")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("median", display_type = "output")`

```{r median, include = F}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(median_life_exp = median(lifeExp))
```




---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("twosummaries")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("twosummaries", display_type = "output")`

```{r twosummaries, include = F}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  summarise(median_life_exp = median(lifeExp),
            mean_life_exp = mean(lifeExp))
```




---
class: center, inverse, middle

# Wrangle to it. #`r i`
---

## Target #`r i` `r i <- i+1`

```{r, echo = F, eval = T, code = knitr::knit_code$get("overall")}
```

---
class: center, middle

### Walk through

---

`r chunk_reveal("overall", display_type = "output")`

```{r overall, include = F}
gapminder %>% 
  filter(year == 2002) %>% 
  summarise(median_life_exp = median(lifeExp),
            mean_life_exp = mean(lifeExp))
```
